---
title: "01-exercises"
author: "Mark Cooke"
date: "April 11, 2016"
output: html_document
---


## Exercise 1:

Write a function (f) that takes a vector of numbers, **x**, and returns a vector of numbers such that each element containing the product of every element of x except the element of x with the same index. 


### Example 

    > x <- c( 1, 5, 2, 8 ) 
    > f(x) 
    [1] 80 16 40 10
    # 5*2*8, 1*2*8, 1*5*8, 1*2*5  


### Solution 

```{r}
f <- function(x) {
  y<-prod(x)/x # may have NaN at indices of zeros in x
  z<-which(x%in%0) # get indices of zeros in x
  lz<-length(z)
  if(lz>1){ # if > 1 zeros, all NaN = 0
    for(idx in z){
      y[idx]<-0
    }
  }else if (lz==1){ # if 1 zero, NaN = prod(other elements)
    xmod<-x[-z]
    y[z]<-prod(xmod)
  }
  return(y)
}
```


## Exercise 2

Write a function f(x) to accept an integer vector, and returns a vector with those numbers … except for: 
   multiples of 3 = “Fizz” 
   multiples of 5 = “Buzz”. 
   multiples of 3 and 5 = “FizzBuzz”
   
### Example

    > x <- 1:20
    > f(1:20) 
    # 1 2 Fizz 4 Buzz Fizz 7 8 Fizz Buzz 11 Fizz 13 14 FizzBuzz 16 17 Fizz 19 Buzz


### Solution

```{r}
f <- function(x) {
  mulThree<-which((x%%3)%in%0)
  mulFive<-which((x%%5)%in%0)
  mulFifteen<-which((x%%15)%in%0)
  for(idx in mulThree){
    x[idx]<-"Fizz"
  }
  for(idx in mulFive){
    x[idx]<-"Buzz"
  }
  for(idx in mulFifteen){
    x[idx]<-"FizzBuzz"
  }
  return(x)
}
```
